Analyzing Trends in U.S. Residential Real Estate Prices

### 2024-04-17 outputs/paper/paper.qmd
paper updated

# load packages
packages <- c("tidyverse", "ggplot2", "janitor", "readr", "knitr", "modelsummary", "testthat", "kableExtra", "viridis", "lubridate", "maps", "mgcv")

@@

# read in data
all_house_price <- read_csv("../../inputs/data/all_house_price.csv")
house_prices <- read_csv("../../outputs/data/merged_house_price.csv")

@@

Data used in this paper was cleaned, processed and tested with the programming language R [@citeR]. Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `readr` [@citeReadr], `knitr` [@citeKnitr], `modelsummary` [@citeModelsummary], `testthat` [@citeTestthat], `KableExtra` [@citeKableEx], `viridis` [@citeViridis], `lubridate` [@citeLubridate], `maps` [@citeMaps], `mgcv` [@citeMgcv].

@@

The results of the Generalized Additive Model (GAM) as presented in Table 6 show a significant positive intercept, mirroring the polynomial regression model, which suggests a high base price for houses at the outset of the analysis. The R-squared value of 0.506 represents an improvement over the linear models, explaining approximately 50.6% of the variability in house prices. This indicates the GAM's enhanced ability to capture the variance in house prices through its non-linear approach.

@@

The two charts in @fig-GAM_fit each depict the smoothed relationship between the predictor and the response variable (house price), with a solid line representing the estimated function and dashed lines indicating the confidence intervals around the estimate.

>>>>

# load packages
packages <- c("tidyverse", "ggplot2", "janitor", "readr", "knitr", "modelsummary", "testthat", "kableExtra", "viridis", "lubridate", "maps", "mgcv", "arrow")

@@

# read in data
all_house_price <- read_csv("../../inputs/data/all_house_price.csv")
house_prices <- read_parquet("../../outputs/data/merged_house_price.parquet")

@@

Data used in this paper was cleaned, processed and tested with the programming language R [@citeR]. Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `readr` [@citeReadr], `knitr` [@citeKnitr], `modelsummary` [@citeModelsummary], `testthat` [@citeTestthat], `KableExtra` [@citeKableEx], `viridis` [@citeViridis], `lubridate` [@citeLubridate], `maps` [@citeMaps], `mgcv` [@citeMgcv], `Arrow` [@citeArrow].

@@

The results of the Generalized Additive Model (GAM) as presented in @tbl-nonlinear_model_result show a significant positive intercept, mirroring the polynomial regression model, which suggests a high base price for houses at the outset of the analysis. The R-squared value of 0.506 represents an improvement over the linear models, explaining approximately 50.6% of the variability in house prices. This indicates the GAM's enhanced ability to capture the variance in house prices through its non-linear approach.

@@

In @fig-GAM_fit, the numbers in the parentheses, such as 10.83 and 2.74, represent the smoothing parameters or effective degrees of freedom. These parameters control the flexibility of the spline, balancing the fit of the model against the risk of overfitting by adjusting the "wiggliness" of the curve. The values on the y-axis, which range from negative to positive, quantify the contribution of each smooth term to the predicted value of the response variable, such as house prices. For example, these y-axis values illustrate how changes in the year or the number of bedrooms influence the predicted house price, relative to a certain baseline.

The solid line depicted in the plot is the estimated smooth curve, which shows the modeled relationship between the predictor and the response variable. The dashed lines surrounding the solid line typically represent confidence intervals, providing a range within which the true effect is likely to lie with a certain level of confidence. This visualization helps in understanding the certainty of the model predictions and the potential variability in the effect of each predictor on the response variable.
